---
title: "concRete.rmd"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "hide", fig.width=8, fig.height=4)
options(scipen = 0, digits = 3)  # controls base R output
# check if you have ISLR package, if not, install it
if(!require('pacman')) {install.packages('pacman')}
pacman::p_load(ISLR, readxl, tidyverse, magrittr, dplyr, ggplot2, ggrepel)
```

```{r some EDA stuffs, include=FALSE}
solar_gen_weather <- read.csv("data/solar-gen-weather.csv")
str(solar_gen_weather)
head(solar_gen_weather)
summary(solar_gen_weather)

# we remove angle_of_incidence because this depends on the tilt of the PV panel and not on the geographic placement
solar_gen_weather <- solar_gen_weather %>% select(-angle_of_incidence)


fit <- lm(generated_power_kw ~ ., solar_gen_weather)
fit <- update(fit, .~. - medium_cloud_cover_mid_cld_lay)
fit <- update(fit, .~. - wind_direction_900_mb)
fit <- update(fit, .~. - wind_direction_10_m_above_gnd)
fit <- update(fit, .~. - wind_speed_80_m_above_gnd)
fit <- update(fit, .~. - total_precipitation_sfc)
fit <- update(fit, .~. - snowfall_amount_sfc)
fit <- update(fit, .~. - wind_speed_10_m_above_gnd)
fit <- update(fit, .~. - wind_speed_900_mb)

summary(fit)

# graph <- solar_gen_weather %>% ggplot(aes(x = shortwave_radiation_backwards_sfc, y = generated_power_kw)) + geom_point() + geom_smooth(method="lm")
# graph
# 
# sgw <- solar_gen_weather %>% rename(temp = temperature_2_m_above_gnd, humidity = relative_humidity_2_m_above_gnd, )

```

After doing backwards elimination we find that the most important variables are as follows:
temperature_2_m_above_gnd: measured in degrees celsius
relative_humidity_2_m_above_gnd: measured in percent
mean_sea_level_pressure_MSL: measured in millibars
total_cloud_cover_sfc
high_cloud_cover_high_cld_lay
low_cloud_cover_low_cld_lay
shortwave_radiation_backwards_sfc
wind_direction_80_m_above_gnd
wind_gust_10_m_above_gnd
zenith: calculated based on time of day and longitude and latitude
azimuth: calculated based on time of day and longitude and latitude

